---
title: "FPLab3"
Author: "Janette Avelar David Fainstein Joe Swinehart Makayla Whitney"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(purrr)
library(tidyverse)
```

## Part A: Foundations
**1. Briefly name and describe the three fundamental components of a function.**

the body(), the actual code inside the function, where you can see what's going on behind the curtain. For a cheap example of this, #body(body) works!

the formals(), the arguments which are supplied the function. Fun fact...if you try to get formals for a primitive function, you get a NULL response.

the environment(), the “map” of the location of the function’s variables. This is largely a mystery to us at this point.

**2. Describe three different ways functions can be stored/applied and an example of when you might want to use each version.**

Typically, functions are stored as objects by assigning them to a name. We do this the same way we store any other object in the R environment; using the `<-` operator. This can be considered the "go-to" storage for a function, after which we apply it as we would any other function: calling the function name, and passing necessary arguments within `()`.

You don't have to store functions as objects, however. In some cases, it's beneficial to use anonymous functions that are not bound to names. To do this, we specify the function we want to use within a loop without assigning it to a name. Once it's run, it's done.

The final way to save functions is to store them within a list. You access them as you would any other object within a list, by using subset operators like `$`, `[]`, or `[[]]`.


## Part B: Applied practice

**1. Write a function to calculate the mean that removes missing data before conducting the calculation. You may not use base::mean or any similar function that conducts a mean calculation. Include a warning in your function if missing data have been removed that prints the total number of cases that were removed.**

```{r part b1}
naless_mean <- function(x) {
  if(any(is.na(x))) {
    y <- na.omit(x)
    z <- length(is.na(x))
    a <- sum(y)/length(y)
    warning("Missing values were found:", z, a)
  }
sum(x)/length(x)
}

```


**2. Test your function to make sure it (a) provides the expected results, and (b) gives identical output to base::mean when na.rm = TRUE. Make sure that you test your data against a vector that has missing data.**

```{r part b2}
naless_mean(mtcars$mpg)
mean(mtcars$mpg)

```


**3. Turn the following three lines of code into three different functions. Make sure to give them meaningful names. Test the functions to make sure they provide the expected output.**

```{r part b3}
# Count number of missing elements in a vector
sum(is.na(airquality$Ozone))

#function to count missing elements in a vector
missing_els <- function(x){
  mean(is.na(x))
}

#testing missing element function
missing_els(c(rep(NA, 35), rnorm(85)))

# Proportional representation of each level (unique element) in a vector
purrr::map_dbl(split(mtcars$cyl, mtcars$cyl), length) / length(mtcars$cyl)

#function to display proportional representation of each unique element in a vector
proprep_els <- function(cars_v){
  purrr::map_dbl(split(cars_v, cars_v), length) / (length(cars_v))
}

#testing proportional representation function
proprep_els(rep(letters[1:4], each = 25))

# normalize or z-score a vector so the mean is zero and sd is one
(mtcars$mpg - mean(mtcars$mpg, na.rm = TRUE)) / 
  sd(mtcars$mpg, na.rm = TRUE) 

#function to normalize a vector
normalize_v <- function(x) {
  if(any(is.na(x))) {
    total_na <- sum(is.na(x))
    warning(paste(total_na, "observations removed due to missing data"))
    x <- na.omit(x)
  }
  (x - mean(x)) / sd(x)
}

#test function to normalize a vector
test_normalize_v <- cbind(normalize_v(mtcars$mpg),
                          na.omit(scale(mtcars$mpg)))

identical(test_normalize_v[ ,1], test_normalize_v[ ,2])


```

**4. Write a function that takes a data frame as its input and returns a data frame with only the numeric columns.**

```{r part b4}

numericdf <- fucntion(df)

```

**5. Write a function that uses the function you wrote in Question 5, and returns a data frame with the mean and standard deviation of each numeric column. If columns are omitted (because they are non-numeric), print a message stating the specific columns that were not included.**

```{r part b5}

```

